---
title: "Individual 2"
author: "Andrew Sundberg"
date: "April 11, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Lab 5

## Part 1

```{r}
library(ggplot2)
library(tidyverse)
library(dplyr)
acc <- read.csv("https://raw.githubusercontent.com/xdaiISU/ds202materials/master/hwlabs/fars2017/accident.csv", 
                stringsAsFactors = FALSE)
acc$DAY_WEEK = factor(acc$DAY_WEEK)
acc$HOUR = factor(acc$HOUR)
acc$DRUNK_DR = factor(acc$DRUNK_DR)
```


## Day of the week

```{r}
ggplot(acc, aes(x = DAY_WEEK, y = VE_TOTAL)) + geom_boxplot()
```


## Time of day

```{r}
ggplot(acc, aes(x = HOUR, y = VE_TOTAL)) + geom_boxplot()
```


## Drunk Driver

```{r}
ggplot(acc, aes(x = DRUNK_DR, y = VE_TOTAL)) + geom_boxplot()
table(acc$DRUNK_DR)
sum((acc$DRUNK_DR == 1) + (acc$DRUNK_DR == 2) + (acc$DRUNK_DR == 3))
```

The total number of accidents with at least one drunk driver is 8769.


## Part 2

## Filter by person type one

```{r}
person = read.csv("https://raw.githubusercontent.com/xdaiISU/ds202materials/master/hwlabs/fars2017/person.csv")
person1 = person[person$PER_TYP == 1,]
```


## Join the data sets

```{r}
joined = inner_join(acc,person1, by = "ST_CASE")

joined %>% group_by(DAY_WEEK) %>% summarize(num.accidents = n())
joined %>% group_by(HOUR.x) %>% summarize(num.accidents = n())
joined %>% group_by(SEX) %>% summarize(num.accidents = n())

barplot(table(joined$DAY_WEEK), xlab = "Day of the Week", ylab = "Number of Accidents", col = "blue")
barplot(table(joined$HOUR.x), xlab = "Hour of the Day", ylab = "Number of Accidents", col = "red")
barplot(table(joined$SEX), xlab = "Gender", ylab = "Number of Accidents", col = "green")

```

There seem to be more accidents on Friday, Saturday, and Sunday during the week. It looks like the most amount of accidents during the day occur around 5:00 pm to 9:00 pm.  And a majority of people in the accidents are males.


## Part 3

I could not load the glcs data file because it says the vector size is too large.  I worked on this part with my partner
```{r}
#glc = readxl::read_xlsx("C:/Users/Andrew Sundberg/Documents/DS202/ds202_lab5/FRPP GLC United States.xlsx", skip = 1)
#names(glc) = c("Territory",  "State Name", "State Code", "City Code", "City Name", "COUNTY", "County Name", "Country Code", "Old City Name","Date Record Added")

#glc$COUNTY = as.integer(glc$COUNTY)
#p3 = full_join(acc, glc, by = "COUNTY")

#ggplot(p3, aes(x = LONGITUD, y = LATITUDE)) + geom_path(aes(group = COUNTY))
```

